EGGs are adhesive!\\ An algebraic account of equality saturation

### 2025-09-09 jlamp/jlamp.tex
fine

\newcommand{\pbc}{\mathsf{Pb}}
\newcommand{\pbe}{\mathsf{ePb}}
\newcommand{\pbl}{\mathsf{LPb}}
\newcommand{\pble}{\mathsf{eLPb}}
\newcommand{\pbs}{\mathsf{SPb}}
\newcommand{\pbse}{\mathsf{eSPb}}
\usepackage{amsthm,amsmath,amssymb,mathrsfs, dsfont}
%\usepackage{tikz-cd}
\usepackage{hyperref, cleveref} %simboli matematici
\usepackage[all, cmtip]{xy}
\usepackage{thm-restate}

@@

\def\R{\mathsf{R}}
\def\B{\textbf {\textup{B}}}
\def\C{\textbf {\textup{C}}}
\def\D{\textbf {\textup{D}}}
\def\X{\textbf {\textup{X}}}
\def\Y{\textbf {\textup{Y}}}
\def\E{\textbf {\textup{E}}}
\def\T{\textbf {\textup{1}}}
\def\A{\textbf {\textup{A}}}
\def\M{\mathcal{M}}
%categorie varie
\newcommand{\catname}[1]{\textbf{\textup{#1}}}
\newcommand{\lab}{\catname{LHyp}}
\newcommand{\lmo}{\catname{LMH}}
\newcommand{\hyp}{\catname{Hyp}}
\newcommand{\phyp}{\catname{pHyp}}
\newcommand{\pshyp}{\catname{pSep}}
\newcommand{\shyp}{\catname{Sep}}
\newcommand{\hyps}{\catname{Hyp}_{\Sigma}}
\newcommand{\EqHyp}{\catname{Hyp}_{\mathsf{eq}}} %equivalence hypergraphs
\newcommand{\EqsHyp}{\catname{Sep}_{\mathsf{eq}}} 
\newcommand{\EqpHyp}{\catname{pHyp}_{\mathsf{eq}}}
\newcommand{\EqpsHyp}{\catname{pSep}_{\mathsf{eq}}}
\newcommand{\Eqlmo}{\catname{LMH}_{\mathsf{eq}}}
\newcommand{\EqHyps}{\catname{Hyp}_{\mathsf{eq}, \mathcal{G}}}
\newcommand{\EqTG}{\catname{EqTG}}
\newcommand{\EqTGs}{\catname{EqTG}_{\Sigma}}
\newcommand{\GEqTGs}{\catname{GEqTG}_{\Sigma}}
\newcommand{\GEqATGs}{\catname{GEqATG}_{\Sigma}}
\newcommand{\gr}{\textbf{\textup{Graph}}}
\newcommand{\dgr}{\catname{SGraph}}
\newcommand{\dg}{\catname{DAG}}
\newcommand{\rt}{\mathsf{dcl_s}}
\newcommand{\rta}{\mathsf{dcl}}
\newcommand{\rtd}{\mathsf{dcl_{d}}}
\newcommand{\slice}[2]{(\catname{#1}\downarrow{#2})}
\newcommand{\tg}[0]{\catname{TG}_{\Sigma}}
\newcommand{\teg}[0]{\catname{TeGr}_{\Sigma}}
\newcommand{\sv}[0]{\mathsf{Sieves}}
\newcommand{\mono}[1]{\mathsf{Mon}(\catname{#1})}
\newcommand{\mo}[1]{{#1}_\mathsf{Mon}}
\newcommand{\pro}{\mathsf{prod}}
\newcommand{\spro}{\mathsf{ps}}
\newcommand{\prol}{\mathsf{lprod}}
\newcommand{\pred}[1]{{\downarrow}#1}
\newcommand{\colim}[0]{\mathrm{colim}}
\newcommand{\cod}{\mathsf{cod}}
\renewcommand{\sp}{\mathsf{sp}}
\newcommand{\pr}{\mathsf{pr}}
\newcommand{\is}[1]{{#1}_{is}}
\renewcommand{\sup}{\mathsf{sup}}
\newcommand{\cow}[1]{\mathsf{cwd}({#1})}
\renewcommand{\inf}{\mathsf{inf}}
\newcommand{\dom}{\mathsf{dom}}
\newcommand{\dwnarrow}{\downarrow \hspace{-2pt}}
\newcommand{\Dwnarrow}{\Downarrow \hspace{-2pt}}
\newcommand{\egg}{\mathsf{e}\text{-}\catname{EqHyp}}

@@

%questo sparirà, serve a me per non diventare scemo
\setcounter{secnumdepth}{4}
\setcounter{tocdepth}{4}
\tableofcontents

@@

 \begin{prop}
 	Let $M_{\eq}\colon \Eqlmo\to \EqHyp$ and $S_{\eq}\colon \EqsHyp \to \EqHyp$ be the inclusion functors. Let, moreover, $\Gamma_{R(\mathcal{G})}\colon 1\to \hyp$ be the constant functor in $R(\mathcal{G})$. Then
 	\begin{enumerate}
 		\item $\lmo_{\eq, \mathcal{G}}$ is equivalent to $\comma{M_\eq}{\Gamma_{R(\mathcal{G})}}$;
 		\item $\shyp_{\eq, \mathcal{G}}$ is equivalent to $\comma{S_\eq}{\Gamma_{R(\mathcal{G})}}$; 
 		\item $\pshyp_{\mathcal{G}}$ is equivalent to $\pshyp_{\eq}/R(\sp(\pr(\mathcal{G})))$. 
 	\end{enumerate}
 \end{prop}

\begin{defi}We define the following two classes of arrows:
	\begin{itemize}
		\item $\pbl_{\mathcal{G}}$ is the class of all arrows in $\lmo_{\eq, \mathcal{G}}$ sent by $M_{\eq, \mathcal{G}}$ to $\pbc_{\mathcal{G}}$;
		\item $\pbs$ is the class of all arrows in $\pshyp_{\eq, \mathcal{G}}$ sent by the inclusion  to $\pbc_{\mathcal{G}}$.
	\end{itemize}
\end{defi}

@@

\fi 
\section{EGGs}
\label{eggs}
The previous section proved some adhesivity property for the categories $\EqHyp$ and $\EqTG_{\Sigma}$. We extend these results to
encompass equivalence classes which are closed under the target arrow, i.e.~under operator composition for term graphs, thus precisely capturing EGGs.

@@

	We will denote by $\egg$ the full subcategory of $\EqHyp$ whose objects are e-hypergraphs, and by $I\colon \egg \to \EqHyp$ the associated inclusion functor.
\end{defi}

\begin{exa}
	Consider the hypergraph $\mathcal{G}$ of \Cref{exa_2} and consider as quotient the identity $\id{V_\mathcal{G}}\colon V_\mathcal{G}\colon \to\mathcal{G}$. Then the kernel pair of $\id{V^\star_\mathcal{G}} \circ s$ concide with the kernel pair of $s$, which is empty. Thus  $\mathcal{G}$ is, trivially, an e-hypergraph
\end{exa}
A first result that we need is that limits in $\egg$ are computed as in $\EqHyp$. Full proofs are in \Cref{proof:elim}.

\begin{restatable}{lem}{elim}\label{lem:elim}
	$\egg$ has all limits and $I$ creates them.
\end{restatable}

\begin{cor}\label{cor:ereg}
	If an arrow $h: \mathcal{G \to H}$ in $\egg$ is a regular mono in $\egg$ then $I(h)$ is a regular mono in $\hyp$. 
\end{cor}

@@

\noindent
\begin{minipage}[l]{.83\linewidth}
	\begin{restatable}{lem}{epo}\label{lem:epo}
		Suppose that the square on the right is a pushout in $\EqHyp$ and that $m$ is a mono in $\pbc$. If $\mathcal{G}_1$, $\mathcal{G}_2$ and $\mathcal{G}_3$ are e-hypergraphs, then $\mathcal{P}$ is 
		is an e-hypergraph.
		%so too.
		%an object of $\egg$ too.
	\end{restatable}
\end{minipage}\hfill 
\begin{minipage}[r]{.15\linewidth}
	\xymatrix@R=15pt{\mathcal{G}_1 \ar[r]_{h}\ar@{>->}[d]_{m}&\mathcal{G}_2\ar[d]^{n}\\\mathcal{G}_3\ar[r]^{z}&\mathcal{P}}
\end{minipage}

Let now $\pbe$ be the class of arrows in $\egg$ that are sent to $\pbc$ by $I$. By definition and \Cref{lem:pbmono} we have that such class is closed under composition and decomposition, it contains all isomorphisms and it is stable under pullbacks. Using \Cref{lem:pbmono} and \Cref{lem:epo} we can further deduce that it is stable under pushouts. Then \Cref{thm:slice-functors} and \Cref{cor:eqade} yield the following.

@@

\subsection{E-term graphs, a.k.a. EGGs}

We now turn our attention to the labelled context.

\begin{defi} We say that an object $(\mathcal{H}, l)$ of $\EqHyps$ is a \emph{labelled e-hypergraph} if $\mathcal{H}$ is an e-hypergraph. We define the category $\egg_\Sigma$ as the full subcategory of $\EqHyps$ given by labelled e-hypergraphs. We denote by $Z_\Sigma$ the corresponding inclusion functor .
\end{defi}

To prove some adhesivity property of $\egg_\Sigma$, we begin with the following elementary, yet useful observation.

\begin{rem}\label{rem:algegg}
	Given a signature $\Sigma = (O_\Sigma, \ari_\Sigma)$,
	the hypergraph 
	$R(\mathcal{G}^\Sigma) = (O_\Sigma, 1, 1, \ari_\Sigma, \gamma_1, \id{1})$
	is an object of $\egg$.
	Indeed, under the identification of $1^\star$ with $\mathbb{N}$, the kernel $(S, \pi_1, \pi_2)$  of $\id{\mathbb{N}} \circ \ari_\Sigma$, is given by
	$S:=\{(o_1, o_2)\in O_\Sigma \times O_\Sigma \mid \ari_\Sigma(o_1)=\ari_\Sigma(o_2)\}$
	equipped with the two projections.  On the other hand, both $\id{\mathbb{N}} \circ \gamma_1 \circ \pi_1$ and $\id{\mathbb{N}}\circ \gamma_1 \circ \pi_2$ are the function $O_\Sigma \to \mathbb{N}$ constant in $1$.
\end{rem}

\Cref{rem:algegg} now implies at once the following result.

\begin{prop}\label{prop:varie}
	Let $\Sigma$ be an algebraic signature. Then the following facts hold
	\begin{enumerate}
		\item $\egg_\Sigma$ is equivalent to $\egg/R(\mathcal{G}^\Sigma)$;
		\item there exists a functor $W_\Sigma \colon \egg_\Sigma\to \egg$ forgetting the labeling which creates all colimits, pullbacks and equalizers. 
	\end{enumerate}
\end{prop}

Let $\pbe_\Sigma$ be the class of morphisms in $\egg_\Sigma$ whose image in $\egg$ lies in $\pbe$. Notice that $\pbe_\Sigma$ is also the class of arrows whose image through $Z_\Sigma$ is in $\pbc_\Sigma$.
%
By \Cref{thm:slice-functors} we get the following result.

\begin{cor}\label{cor:eggade1}
	$\egg_\Sigma$ is $\pbe_\Sigma$-adhesive.
\end{cor}

We turn now to term graphs with equivalence.

@@

\begin{rem}
	By definition we also have an inclusion functor $Y_\Sigma\colon \eg\to \egg_\Sigma$.
\end{rem}

\noindent 
\begin{minipage}[l]{.58\linewidth}
	\begin{rem}
		Now that we have put all the structures of this work in place, it is worthwhile to give a visual map of all the categories that we have discussed/introduced and of the relationships between them. We will use the curved arrows to denote full and faithful inclusions.
	\end{rem}
\end{minipage}\hfill
\begin{minipage}[r]{.20\linewidth}
	\xymatrix@R=10pt@C=20pt{\eg \ar@{^{(}->}[r]^-{Y_\Sigma} \ar@{^{(}->}[d]_{K_\Sigma} & \egg_\Sigma \ar@{^{(}->}[d]^{Z_\Sigma} \ar[r]^{W_\Sigma}& \egg \ar@{^{(}->}[d]^{I}\\ \EqTG_{\Sigma} \ar@{^{(}->}[r]^-{J_\Sigma}\ar[d]_{S_\Sigma}& \EqHyps \ar[r]^-{V_\Sigma} \ar[d]_{T_\Sigma}& \EqHyp \ar[d]^{T}\\ \tg \ar@{^{(}->}[r]^-{I_\Sigma}& \hyp_{\Sigma} \ar[r]^-{U_\Sigma}  &\hyp}
\end{minipage}

We can now prove our last three results. The reader can find the proof in \Cref{proof:tlim}.

@@

\begin{cor}
$\eg$ is $\mathcal{T}_\Sigma$-adhesive.
\end{cor}

\subsubsection{BOH}\label{sec:sepegg}

qui vanno gli egg fatti con i separati

\subsubsection{BOH}\label{sec:lmegg}

qui vanno gli egg fatti con i left monogamous (probabilmente sussume la 5.2)

@@

\noindent
\begin{minipage}[l]{.8\linewidth}
\begin{lem}\label{lem:po_set}
Suppose that the square aside is a pushout, with $m\colon A\mto B$ an injection. Let $\iota_{m}\colon E\mto B$ be the inclusion of the complement of the image of $m$. Then $(D, \{n, h\circ \iota_m \})$ is a coproduct. In particular, $h\circ \iota_m$ is mono.
\end{lem}
\end{minipage}
\hfill 
\begin{minipage}[r]{.23\linewidth}
\xymatrix@R=15pt@C=30pt{A \ar@{>->}[r]^{m}  \ar[d]_{g}& B \ar[d]^{h}\\ C \ar@{>->}[r]_{n} & D}
\end{minipage}
\begin{proof}
		 Let $f\colon E\to Z$ and $k\colon C\to Z$ be two arrows with the same codomain. By definition of complement and since $m$ is mono $(B, \{m, \iota_m\})$ is a coproduct. Let $\phi\colon B\to Z$ be the unique arrow such that \[\phi\circ m=k\circ g \qquad \phi \circ \iota_m=f\] 
		By the universality of pushouts there is a unique arrow $\psi \colon D\to Z$ such that $\psi \circ n=k$ and $\psi \circ h=\phi$, thus 
		\[
		\psi \circ h\circ \iota_m =\phi \circ \iota_m=f\]	
To conclude we have to show that $\psi$ is the unique arrow $D\to Z$ such that $\psi \circ n =k$ and $\psi \circ h\circ \iota_m=f$. Let $\psi'\colon D\to Z$ be another arrow such that $\psi' \circ n=k$ and $\psi' \circ h\circ \iota_m=f$. Then we have
		\[\psi'\circ h\circ m=\psi'\circ n\circ g=k\circ g=\phi \circ m =\psi \circ h\circ m \]
	Since $m$ is mono then $\psi'\circ h=\psi \circ h$,  but $(D, \{n, h\})$ is a pushout and so $\psi'=\psi$. 
\end{proof}

@@

\noindent 
\parbox{11.5cm}{\hspace{15pt} Extensivity means that given a family of objects $\{X_{i}\}_{i\in I}$ and a family of commuting squares as the one on the right, where $j_i$ is a  coprojection, then all the squares are pullbacks if and only if $(Z, \{k_i\}_{i\in I})$ is a coproduct.}\hfill 
\parbox{2cm}{\xymatrix@R=15pt{Z_i \ar[r]_{k_i} \ar[d]_{f_i} & Z\ar[d]^{f} \\ X_i \ar@{>->}[r]^{j_i} & X}}

\noindent
\parbox{11.5cm}{
\begin{rem}\label{rem:disj}
	Notice that extensivity entails that coproducts are \emph{disjoint}, i.e.~the pullback between two coprojections is given by the initial object (with initial maps as coprojections). To see this, let $\{X, \{j_i\}_{i\in I}\}$ be the coproduct of the family $\{X_i\}_{i\in I}$ and $k$ an element of $I$. Then  $(X_k, \{t_i\}_{i\in I})$ 
such that $t_i = \id{X_k}$ if $i=k$, and $?_{X_k}$ otherwise,
	%\end{cases}\]	
	%with
	%\[t_i=\begin{cases}
	%	\id{X_k} & i=k\\?_{X_k} & i\neq k
	%\end{cases}\]
	is a coproduct and so the squares on the right are pullbacks.
\end{rem}}\hfill\parbox{2cm}{\xymatrix@R=15pt{\emptyset \ar@{>->}[d]_{?_{X_i}} \ar@{>->}[r]_{?_{X_k}}& X_k \ar@{>->}[d]^{j_k} \\ X_i \ar@{>->}[r]^{j_i} & X} \xymatrix@R=15pt{X_k \ar[r]_{\id{X}} \ar[d]_{\id{X}}& X_k \ar@{>->}[d]^{j_k}\\X_k \ar@{>->}[r]^{j_k} & X}}

In particular, the previous remark yields at once the following fact.

>>>>

\newcommand{\pbc}{\mathsf{Pb}}
\newcommand{\pbe}{\mathsf{ePb}}
\newcommand{\pbl}{\mathsf{LPb}}
\newcommand{\pble}{\mathsf{eLPb}}
\newcommand{\pbs}{\mathsf{SPb}}
\newcommand{\epbl}{\mathsf{eLPb}}
\newcommand{\epbs}{\mathsf{eSPb}}

\newcommand{\pbse}{\mathsf{eSPb}}
\usepackage{amsthm,amsmath,amssymb,mathrsfs, dsfont}
%\usepackage{tikz-cd}
\usepackage{hyperref, cleveref} %simboli matematici
\usepackage[all, cmtip]{xy}
\usepackage{thm-restate}

@@

\def\R{\mathsf{R}}
\def\B{\textbf {\textup{B}}}
\def\C{\textbf {\textup{C}}}
\def\D{\textbf {\textup{D}}}
\def\X{\textbf {\textup{X}}}
\def\Y{\textbf {\textup{Y}}}
\def\E{\textbf {\textup{E}}}
\def\T{\textbf {\textup{1}}}
\def\A{\textbf {\textup{A}}}
\def\M{\mathcal{M}}
%categorie varie
\newcommand{\catname}[1]{\textbf{\textup{#1}}}
\newcommand{\lab}{\catname{LHyp}}
\newcommand{\lmo}{\catname{LMH}}
\newcommand{\hyp}{\catname{Hyp}}
\newcommand{\phyp}{\catname{pHyp}}
\newcommand{\pshyp}{\catname{pSep}}
\newcommand{\shyp}{\catname{Sep}}
\newcommand{\hyps}{\catname{Hyp}_{\Sigma}}
\newcommand{\EqHyp}{\catname{Hyp}_{\mathsf{eq}}} %equivalence hypergraphs
\newcommand{\EqsHyp}{\catname{Sep}_{\mathsf{eq}}} 
\newcommand{\EqpHyp}{\catname{pHyp}_{\mathsf{eq}}}
\newcommand{\EqpsHyp}{\catname{pSep}_{\mathsf{eq}}}
\newcommand{\Eqlmo}{\catname{LMH}_{\mathsf{eq}}}
\newcommand{\EqHyps}{\catname{Hyp}_{\mathsf{eq}, \mathcal{G}}}
\newcommand{\EqTG}{\catname{EqTG}}
\newcommand{\EqTGs}{\catname{EqTG}_{\Sigma}}
\newcommand{\GEqTGs}{\catname{GEqTG}_{\Sigma}}
\newcommand{\GEqATGs}{\catname{GEqATG}_{\Sigma}}
\newcommand{\gr}{\textbf{\textup{Graph}}}
\newcommand{\dgr}{\catname{SGraph}}
\newcommand{\dg}{\catname{DAG}}
\newcommand{\rt}{\mathsf{dcl_s}}
\newcommand{\rta}{\mathsf{dcl}}
\newcommand{\rtd}{\mathsf{dcl_{d}}}
\newcommand{\slice}[2]{(\catname{#1}\downarrow{#2})}
\newcommand{\tg}[0]{\catname{TG}_{\Sigma}}
\newcommand{\teg}[0]{\catname{TeGr}_{\Sigma}}
\newcommand{\sv}[0]{\mathsf{Sieves}}
\newcommand{\mono}[1]{\mathsf{Mon}(\catname{#1})}
\newcommand{\mo}[1]{{#1}_\mathsf{Mon}}
\newcommand{\pro}{\mathsf{prod}}
\newcommand{\spro}{\mathsf{ps}}
\newcommand{\prol}{\mathsf{lprod}}
\newcommand{\pred}[1]{{\downarrow}#1}
\newcommand{\colim}[0]{\mathrm{colim}}
\newcommand{\cod}{\mathsf{cod}}
\renewcommand{\sp}{\mathsf{sp}}
\newcommand{\pr}{\mathsf{pr}}
\newcommand{\is}[1]{{#1}_{is}}
\renewcommand{\sup}{\mathsf{sup}}
\newcommand{\cow}[1]{\mathsf{cwd}({#1})}
\renewcommand{\inf}{\mathsf{inf}}
\newcommand{\dom}{\mathsf{dom}}
\newcommand{\dwnarrow}{\downarrow \hspace{-2pt}}
\newcommand{\Dwnarrow}{\Downarrow \hspace{-2pt}}
\newcommand{\egg}{\mathsf{e}\text{-}\catname{Hyp}_\eq}
\newcommand{\eggl}{\mathsf{e}\text{-}\catname{Hyp}_{\eq, \mathcal{G}}}
\newcommand{\elm}{\mathsf{e}\text{-}\catname{LMO}_\eq}
\newcommand{\eps}{\mathsf{e}\text{-}\catname{pSep}_\eq}
\newcommand{\lelm}{\mathsf{e}\text{-}\catname{LMO}_{\eq, \mathcal{G}}}
\newcommand{\leps}{\mathsf{e}\text{-}\catname{pSep}_{\eq, \mathcal{G}}}

@@

%questo sparirà, serve a me per non diventare scemo
%\setcounter{secnumdepth}{4}
%\setcounter{tocdepth}{4}
%\tableofcontents

@@

 \begin{prop}
 	Let $M_{\eq}\colon \Eqlmo\to \EqHyp$ and $S_{\eq}\colon \EqsHyp \to \EqHyp$ be the inclusion functors. Let, moreover, $\Gamma_{R(\mathcal{G})}\colon 1\to \EqHyp$ be the constant functor in $R(\mathcal{G})$. Then
 	\begin{enumerate}
 		\item $\lmo_{\eq, \mathcal{G}}$ is equivalent to $\comma{M_\eq}{\Gamma_{R(\mathcal{G})}}$;
 		\item $\shyp_{\eq, \mathcal{G}}$ is equivalent to $\comma{S_\eq}{\Gamma_{R(\mathcal{G})}}$; 
 		\item $\pshyp_{\mathcal{G}}$ is equivalent to $\pshyp_{\eq}/R(\sp(\pr(\mathcal{G})))$. 
 	\end{enumerate}
 \end{prop}

\begin{defi}We define the following two classes of arrows:
	\begin{itemize}
		\item $\pbl_{\mathcal{G}}$ is the class of all arrows in $\lmo_{\eq, \mathcal{G}}$ sent by $M_{\eq, \mathcal{G}}$ to $\pbc_{\mathcal{G}}$;
		\item $\pbs_{\mathcal{G}}$ is the class of all arrows in $\pshyp_{\eq, \mathcal{G}}$ sent by the inclusion  to $\pbc_{\mathcal{G}}$.
	\end{itemize}
\end{defi}

@@

\fi 
\section{EGGs}
\label{eggs}
The previous section proved some adhesivity property for the categories $\EqHyp$, $\Eqlmo$,  $\EqpsHyp$ and for their labelled variants. We extend these results to
encompass equivalence classes which are closed under the target arrow, i.e.~under operator composition for term graphs, thus precisely capturing EGGs.

@@

	We will denote by $\egg$ the full subcategory of $\EqHyp$ whose objects are e-hypergraphs, and by $E\colon \egg \to \EqHyp$ the associated inclusion functor.
\end{defi}

\begin{exa}
	Consider the hypergraph $\mathcal{G}$ of \Cref{exa_2} and take as quotient the identity $\id{V_\mathcal{G}}\colon V_\mathcal{G}\colon \to\mathcal{G}$. Then the kernel pair of $\id{V^\star_\mathcal{G}} \circ s$ concide with the kernel pair of $s$, which is empty. Thus  $\mathcal{G}$ is, trivially, an e-hypergraph
\end{exa}

A first result that we need is that limits in $\egg$ are computed as in $\EqHyp$. Full proofs are in \Cref{proof:elim}.

\begin{restatable}{lem}{elim}\label{lem:elim}
	$\egg$ has all limits and $E$ creates them.
\end{restatable}

\begin{cor}\label{cor:ereg}
	If an arrow $h: \mathcal{G \to H}$ in $\egg$ is a regular mono in $\egg$ then $E(h)$ is a regular mono in $\hyp$. 
\end{cor}

@@

\noindent
\begin{minipage}[l]{.83\linewidth}
	\begin{restatable}{lem}{epo}\label{lem:epo}
		Suppose that the square on the right is a pushout in $\EqHyp$ and that $m$ is a mono in $\pbc$. If $\mathcal{G}_1$, $\mathcal{G}_2$ and $\mathcal{G}_3$ are e-hypergraphs, then $\mathcal{P}$ is 
		is an e-hypergraph.
		%so too.
		%an object of $\egg$ too.
	\end{restatable}
\end{minipage}\hfill 
\begin{minipage}[r]{.15\linewidth}
	\xymatrix@R=10pt{\mathcal{G}_1 \ar[r]_{h}\ar@{>->}[d]_{m}&\mathcal{G}_2\ar[d]^{n}\\\mathcal{G}_3\ar[r]^{z}&\mathcal{P}}
\end{minipage}

Let now $\pbe$ be the class of arrows in $\egg$ that are sent to $\pbc$ by $E$. By definition and \Cref{lem:pbmono} we have that such class is closed under composition and decomposition, it contains all isomorphisms and it is stable under pullbacks. Using \Cref{lem:pbmono} and \Cref{lem:epo} we can further deduce that it is stable under pushouts. Then \Cref{thm:slice-functors} and \Cref{cor:eqade} yield the following.

@@

As in the previous section, we can focus ourselves to two subcategories of $\egg$.

\begin{defi}We will define $\elm$, $\eps$ as the full subcategories of $\egg$ given by those e-hypergraphs which are, respectively, left-monogamous and pruned and separated. We will denote by $M_{\mathsf{e}}$ and by $S_{\mathsf{e}}$ the corresponding functors.
\end{defi}

Let $\epbl$ and $\epbs$ be the classes of arrows in  $\elm$, $\eps$ which lies, respectively in $\pbl$ and $\pbs$. By Lemmas \ref{lem:equ}, \ref{lem:coequ} and \Cref{cor:pbe} we obtain immediately the following result.

\begin{cor}
	$\elm$ is $\epbl$-adhesive, while $\eps$ is $\epbs$-adhesive.
\end{cor}

\subsection{Labelled e-hypergraphs}\label{sec:sepegg}

Finally, we can turn our attention to the labelled context.

\begin{defi} Let $\mathcal{G}$ be an hypergraph. We say that an object $(\mathcal{H}, (f,g))$ of $\EqHyps$ is a \emph{labelled e-hypergraph} if $\mathcal{H}$ is an e-hypergraph. We define the category $\eggl$ as the full subcategory of $\EqHyps$ given by labelled e-hypergraphs. We denote by $E_\mathcal{G}$ the corresponding inclusion functor .
\end{defi}

From the definitions, we get at once the follosing result.

\begin{prop}\label{prop:varie}
	Let $\mathcal{G}$ be an hypergraph and $\Gamma_{R(\mathcal{G})}$ the costant functor in $R(\mathcal{G})$. Then the following facts hold
	\begin{enumerate}
		\item $\eggl$ is equivalent to $\comma{E}/\Gamma_{R(\mathcal{G})}$;
		\item there exists a functor $W_\mathcal{G}\colon \eggl\to \egg$ forgetting the labeling which creates all colimits, pullbacks and equalizers. 
	\end{enumerate}
\end{prop}

Let $\pbe_\mathcal{G}$ be the class of morphisms in $\eggl$ whose image in $\egg$ lies in $\pbe$. Notice that $\pbe_\mathcal{G}$ is also the class of arrows whose image through $E_\mathcal{G}$ is in $\pbc_\mathcal{G}$.
%
By \Cref{thm:slice-functors} we get the following result.

\begin{cor}\label{cor:eggade1}
	$\eggl$ is $\pbe_\mathcal{G}$-adhesive.
\end{cor}

Now, let $\lelm$ and $\leps$ be the full subcategories of $\eggl$ given by the objects whose underlying hypergraphs with equivalence are, respectively, left-monogamous and pruned and separated.

\begin{rem}
When $\mathcal{G}$ is the hypergraph associated to an algebraic signature, the object of  $\lelm$ coincides with what in \cite{uovadesive} are called \emph{e-termgraphs with equivalence}. 
\end{rem}

\begin{lem}Let $\mathcal{G}$ be an hypergraph and $\Gamma_{R(\mathcal{G})}$ the costant functor in $R(\mathcal{G})$. 
	\begin{enumerate}
		\item $\lelm$ is equivalent to $\comma{E\circ M_{\mathsf{e}}}{\Gamma_{R(\mathcal{G})}}$
		\item $\leps$ is equivalent to $\comma{E\circ S_{\mathsf{e}}}{\Gamma_{R(\mathcal{G})}}$
	\end{enumerate}
\end{lem}

Define $\epbl_{\mathcal{G}}$ and $\epbs_{\mathcal{G}}$ as the classes of arrows whose underlying morphisms are in $\epbl$ or in $\epbs$. The following result now follows at once.

\begin{cor}
	$\lelm$ is $\epbl_{\mathcal{G}}$-adhesive, while $\leps$ is $\epbs_{\mathcal{G}}$-adhesive.
\end{cor}

\iffalse 
We turn now to term graphs with equivalence.

@@

\begin{rem}
	By definition we also have an inclusion functor $Y_\Sigma\colon \eg\to {\egg}_\Sigma$.
\end{rem}

\noindent 
\begin{minipage}[l]{.58\linewidth}
	\begin{rem}
		Now that we have put all the structures of this work in place, it is worthwhile to give a visual map of all the categories that we have discussed/introduced and of the relationships between them. We will use the curved arrows to denote full and faithful inclusions.
	\end{rem}
\end{minipage}\hfill
\iffalse \begin{minipage}[r]{.20\linewidth}
	\xymatrix@R=10pt@C=20pt{\eg \ar@{^{(}->}[r]^-{Y_\Sigma} \ar@{^{(}->}[d]_{K_\Sigma} & \egg_\Sigma \ar@{^{(}->}[d]^{Z_\Sigma} \ar[r]^{W_\Sigma}& \egg \ar@{^{(}->}[d]^{I}\\ \EqTG_{\Sigma} \ar@{^{(}->}[r]^-{J_\Sigma}\ar[d]_{S_\Sigma}& \EqHyps \ar[r]^-{V_\Sigma} \ar[d]_{T_\Sigma}& \EqHyp \ar[d]^{T}\\ \tg \ar@{^{(}->}[r]^-{I_\Sigma}& \hyp_{\Sigma} \ar[r]^-{U_\Sigma}  &\hyp}
\end{minipage}
\fi 
We can now prove our last three results. The reader can find the proof in \Cref{proof:tlim}.

@@

\begin{cor}
$\eg$ is $\mathcal{T}_\Sigma$-adhesive.
\end{cor}
\fi

@@

\noindent
\begin{minipage}[l]{.85\linewidth}
\begin{lem}\label{lem:po_set}
Suppose that the square aside is a pushout, with $m\colon A\mto B$ an injection. Let $\iota_{m}\colon E\mto B$ be the inclusion of the complement of the image of $m$. Then $(D, \{n, h\circ \iota_m \})$ is a coproduct. In particular, $h\circ \iota_m$ is mono.
\end{lem}
\end{minipage}
\hfill 
\begin{minipage}[r]{.23\linewidth}
\xymatrix@R=15pt@C=30pt{A \ar@{>->}[r]_{m}  \ar[d]_{g}& B \ar[d]^{h}\\ C \ar@{>->}[r]^{n} & D}
\end{minipage}
\begin{proof}
		 Let $f\colon E\to Z$ and $k\colon C\to Z$ be two arrows with the same codomain. By definition of complement and since $m$ is mono $(B, \{m, \iota_m\})$ is a coproduct. Let $\phi\colon B\to Z$ be the unique arrow such that \[\phi\circ m=k\circ g \qquad \phi \circ \iota_m=f\] 
		By the universality of pushouts there is a unique arrow $\psi \colon D\to Z$ such that $\psi \circ n=k$ and $\psi \circ h=\phi$, thus 
		\[
		\psi \circ h\circ \iota_m =\phi \circ \iota_m=f\]	
To conclude we have to show that $\psi$ is the unique arrow $D\to Z$ such that $\psi \circ n =k$ and $\psi \circ h\circ \iota_m=f$. Let $\psi'\colon D\to Z$ be another arrow such that $\psi' \circ n=k$ and $\psi' \circ h\circ \iota_m=f$. Then we have
		\[\psi'\circ h\circ m=\psi'\circ n\circ g=k\circ g=\phi \circ m =\psi \circ h\circ m \]
	Since $m$ is mono then $\psi'\circ h=\psi \circ h$,  but $(D, \{n, h\})$ is a pushout and so $\psi'=\psi$. 
\end{proof}

@@

\noindent 
\begin{minipage}[l]{.83\linewidth}
	\setlength{\parindent}{1.5em}
Extensivity means that given a family of objects $\{X_{i}\}_{i\in I}$ and a family of commuting squares as the one on the right, where $j_i$ is a  coprojection, then all the squares are pullbacks if and only if $(Z, \{k_i\}_{i\in I})$ is a coproduct.
\end{minipage}\hfill
\begin{minipage}[r]{.2\linewidth}
\xymatrix@R=10pt{Z_i \ar[r]_{k_i} \ar[d]_{f_i} & Z\ar[d]^{f} \\ X_i \ar@{>->}[r]^{j_i} & X}
\end{minipage}

\noindent
\begin{minipage}[l]{.83\linewidth}
	\setlength{\parindent}{1.5em}
\begin{rem}\label{rem:disj}
	Notice that extensivity entails that coproducts are \emph{disjoint}, i.e.~the pullback between two coprojections is given by the initial object (with initial maps as coprojections). To see this, let $\{X, \{j_i\}_{i\in I}\}$ be the coproduct of the family $\{X_i\}_{i\in I}$ and $k$ an element of $I$. Consider $(X_k, \{t_i\}_{i\in I})$, where:
	\[X_i=\begin{cases}
		X_k & i=k\\\emptyset  & i\neq k
	\end{cases} \qquad  t_i=\begin{cases}
		\id{X_k} & i=k\\?_{X_k} & i\neq k
	\end{cases}\]
		is a coproduct and so the squares on the right are pullbacks.
	\end{rem}
\end{minipage}\hfill 
\begin{minipage}[r]{.2\linewidth}
	\xymatrix@R=15pt{\emptyset \ar@{>->}[d]_{?_{X_i}} \ar@{>->}[r]^{?_{X_k}}& X_k \ar@{>->}[d]^{j_k} \\ X_i \ar@{>->}[r]_{j_i} & X} \xymatrix@R=15pt{X_k \ar[r]^{\id{X}} \ar[d]_{\id{X}}& X_k \ar@{>->}[d]^{j_k}\\X_k \ar@{>->}[r]_{j_k} & X}
\end{minipage}
\parbox{11.5cm}{}\hfill\parbox{2cm}{}

\noindent 
\begin{minipage}[l]{.83\linewidth}
\begin{prop}\label{prop:isempty}
	Suppose that the square on the right is a pullback. Let $\iota \colon E\mto Y$ be the inclusion of the complement of the image of $p_1$. If there exist two  arrows $t_1\colon T\to E$ and $t_2\colon T\to X$ such that $g\circ \iota\circ t_1= f\circ t_2$ then $T$ is empty.
\end{prop}
\end{minipage}
\hfill
\begin{minipage}[r]{.2\linewidth}
\xymatrix@R=15pt{P \ar[r]_{p_1} \ar[d]_{p_2}& Y \ar[d]^{g}\\X \ar[r]^{f}& Z}
\end{minipage}
\begin{proof}
	By the universal property of pullbacks we get an arrow $t\colon T\to P$ such that
